Supplemental material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-08-09 Manuscript/journal_correspondence/ERL_responseletter.Rmd

**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios. This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration. The manuscript is well written and organized. However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled. And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed. I have provided detailed comments below.**

**L109, why need to assign each cell a state type? What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition. This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur. We have added text and references to this section and the next that better describes the state of each cell (lines xxx).

**L114, transitions between ‘state type’? Is it ‘land cover classes’? According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**  
State type in this model is the unique combination of moisture zone, island, and land cover class, as stated on line 110 of the original manuscript. This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic shrubland on Oʻahu. We have added text that better defines how state type influences transition probabilities ().

**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median? What is the justification that low land use scenario will have rate lower than historical records?**  
To distinguish the impact of a low land use scenario from a high land use scenario on changes in agricultural land area, we sampled rates of agricultural contraction and agricultural expansion in the low land use scenario from a uniform distribution bounded by zero and the minimum historical rate. Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus lowering the cropping intensity of land area over time. We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010). We have added text that justifies bounding agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).

@@

**L140, is state type land cover classes? Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**  
We agree that this was a confusing use of terminology. We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript.

**L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
Again, we agree this was a confusing use of terminology. We have omitted "moisture zone" and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class.

**L162, how litter and dead wood pools were initialized?**  
We have added text stating that litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).

**L161, what is 30-year climate normal come from? Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020? Or is there any bias between 30-year climate normal and RCP projections?**  
The contemporary 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al. 2013, 2014), as depicted in Supplementary figure 3. We have added text that better describes this (supplementary lines xxx-xxx).

@@

**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here. Where is the additional 32 Tg C from? What is ratio of aboveground biomass to total live biomass in IBIS?**  
Asner et al. (2016) only quantifies aboveground forest biomass. The additional 32 Tg C comes from the sum of belowground forest biomass and total live biomass in grasslands, shrublands, and agriculture. We have added text describing the equation we used to estimate belowground forest biomass, which was then used to calibrate the IBIS DGVM (lines xxx-xxx).

**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
We have added text (lines xxx-xxx) listing the range GPP from the calibration dataset (Kimball et al. 2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary Table 1).

@@

**L166-168, is similar multiplier also applied to living biomass? If not, will this single NPP multiplier break the equilibrium between NPP and biomass? For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees. External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**  
We have added maps to the supplementary depicting the spatial distribution of stationary NPP multipliers as well as the spatial distribution of NPP that results from the product of IBIS mean NPP values and the NPP spatial multipliers based on the spatial variability of contemporary climate. In our model, as in the real world, a deline in NPP will result in a decline in biomass accumulation, but the biomass that has already accumulated persists and will not somehow be reduced just because growth rates have slowed. We have added text that better describes how and why we apply the NPP spatial multiplier (xxx).

**Line 173, why assume Rh as ratio of NPP? Is the ratio constant over time and across climate change scenarios? Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks? Because of lack of soil temperature in RCP 4.5 and 8.5? Need more justification for this flux ratio method.** 
The ultimate source of CO~2~ respired from ecosystems is from a single process: photosynthesis. The vast majority of CO~2~ respired from soils was fixed by photosynthesis within a year of being respired, often within hours. A disproportionate fraction of annual variability in ecosystem carbon balance is associated with biophyscial factors (such as temperature and moisture) that modulate photosynthesis rather than respiration. Consequently, biophysical conditions that increase photosynthesis increase respiration over that same time period, and vice versa. We have added text and references to this section justifying the fact that plant productivity and ecosystem respiration are highly correlated on annual timescales, and ecosystem respiration tends to scale with carbon inputs to the soil from photosynthesis (lines xxx). Although we agree that incorporating gridded annual maps of projected future soil moisture and soil temperature could improve the predictive power our model, these data products do not exist for the Hawaiian Islands.

**L177-179, are the growth and decay multipliers varied at annual basis? How are multipliers derived for time periods CMIP5 projections do not cover? Such as period between 2020-2049? It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?** 
Yes, the growth and decay multipliers vary annually. We have added text that describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx). The NPP trend over time under both RCPs is depicted in Figure 3. We believe this is a much more clear and succinct way to convey changes in NPP than in a series of maps of future projected NPP under each of the four scenarios. Presenting projected future NPP as we do in Figure 3 is consistent with the focus of this paper, which is on statewide trends of ecosystem carbon balance under different climate and land use scenarios. We have, however, added a map displaying the spatial variability in NPP used to initiate the LUCAS model (supplementary Figure 5). These spatial trends largely hold true throughout the model run.

**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance. It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall. For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**  
Projections of NPP change by land use and climate change scenario are presented in Figure 3, which we believe best conveys statewide trends of ecosystem carbon fluxes under different climate and land use scenarios. We have added text and figures to the supplementary that better describes the empirical NPP model (supplementary lines xxx and supplementary Figure 5).

**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript. We have added text to the supplementary climate section that provides more detail on climate projections (supplementary lines xxx).

**L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ to 7.95 Tg y^-1^), and the ranges of statewide annual NPP estimates  broadly overlapped for each of these years(2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^). This modeled slight decline in NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change. We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the discussion section (lines xxx-xxx).

**L242. It is misleading to state ‘climate change led to increased Rh over time’ as figure 3 shows decreasing trend of Rh under high land use scenario and two RCP climate change scenarios. It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
We agree, this phrasing was misleading. We have altered the text as suggested by this referee (lines xxx).

**L244, why the increase in total soil carbon stock (figure 2) will not enhance Rh? Is it due to the assumption made by this work that Rh is only ratio of NPP and latter decrease over time? What are drivers of Rh decrease?**  
Soil carbon stocks increase consistently only under the RCP 4.5 radiative forcing scenario. Under RCP 8.5, soil carbon stocks reach an asymptote by 2080 in the low land use scenario and begin to decline slightly in the high land use scenario. Also, Rh only declines consistently under the high land use scenario (Figure 3). In the low land use scenario Rh remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5. As stated previously, NPP is a direct driver of Rh because the vast majority of carbon respired heterotrphically from terrestrial ecosystems was fixed by photosynthesis that same year (Kuzyakov et al. 2010; see lines xxx of revised manuscript).

**L244, why Rh shows very sharp decrease before 2020 under all land use scenarios with all RCPs, such decrease rate seems to be higher than the rest period. Are these decreases supported by observations?**  
Land use change rates are the same in all four scenarios from 2010-2020, which explains why Rh behaves consistently across these scenarios. We have added text to make that clear in the Methods section (lines xxx). The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 5). We have added text that describes these trends (lines xxx).

@@

**The manuscript entitled “Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change” is interesting and suitable for publication but several mistakes in the manuscript should be corrected before acceptance.**

**Specific comments**

**Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
We have carefully checked the manuscript for typographical errors and corrected them on lines x, y, and z of the revised manuscript. We have also confirmed that all acronyms are spelled out upon initial use.

**All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
We have increased the resolution of all figures to 400 dpi.

### 2021-08-09 Manuscript/HI_Model_ms.Rmd
updated manuscript files, increased resolution to 400 dpi for all figure files

Initial carbon stocks and baseline carbon flows were derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017. We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017]. Mean annual GPP from @kimball_evaluating_2017 ranged from 0.123 - 3.88 kg C m^-2^ y^-1^ across the Hawaiian Islands, resulting in a NPP-to-GPP ratio ranging from 0.37 - 0.56.

We initiated net primary production for each simulation cell in LUCAS as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019]. The spatial NPP multipliers were the NPP anomaly for each simulation cell relative to mean empirically-derived NPP values for each state type (supplemental Figure 4). Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each state type to the microclimate-adjusted NPP value for each simulation cell. All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].

>>>>

Initial carbon stocks and baseline carbon flows were derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] that was converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017. We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017]. Mean annual GPP from @kimball_evaluating_2017 ranged from 0.123 - 3.88 kg C m^-2^ y^-1^ across the Hawaiian Islands, resulting in a NPP-to-GPP ratio ranging from 0.37 - 0.56.

We initiated net primary production in LUCAS for each simulation cell as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019]. The spatial NPP multipliers were the NPP anomaly for each simulation cell relative to mean empirically-derived NPP values for each state type (supplemental Figure 4). Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each state type to the microclimate-adjusted NPP value for each simulation cell. All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].

### 2021-08-09 Manuscript/HI_Model_supp.Rmd
updated manuscript files, increased resolution to 400 dpi for all figure files

Initial values for net primary production (NPP) in LUCAS were calculated using a lookup table derived from the Integrated Biosphere Simulator (IBIS; supplemental Table 1). These values were adjusted with spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate (supplemental Figure 4) as described in detail in @sleeter_effects_2018 and @sleeter_effects_2019. Briefly, we derived the set of NPP spatial multipliers by first calculating NPP independently of the IBIS values for each simulation cell using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] based on Hawaiʻi-specific climate data (supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014). We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each state type (i.e., the unique combination of moisture zone and land cover class; supplemental Figure 4). These NPP multipliers were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4). 

>>>>

Initial values for net primary production (NPP) in LUCAS were calculated using a lookup table derived from the Integrated Biosphere Simulator (IBIS; supplemental Table 1). These values were adjusted with spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate (supplemental Figure 4) as described in detail in @sleeter_effects_2018 and @sleeter_effects_2019. Briefly, we derived the set of NPP spatial multipliers by first calculating NPP independently of the IBIS values for each simulation cell using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] based on Hawaiʻi-specific climate data (supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014). We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each state type (i.e., the unique combination of moisture zone and land cover class; supplemental Figure 4). These NPP multipliers (range = 0.27 - 5.2) were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4). 
